---
title: "Notebook"
output: github_document
---


```{r setup, message=FALSE, include=FALSE}
library(tidyverse)
theme_set(theme_light())
```

### Define data

```{r}
time_stretch <- 4
time_shift <- 5
exp_stretch <- 6
exp_shift <- 4
```

```{r}
data <- data.frame(
  locus_name = "A_1",
  accession = "Col0",
  tissue = "apex",
  timepoint = c(10L, 13L, 16L, 19L, 
                21L, 24L),
  mean_cpm = c(
    10.3020734833646, 8.1765718166656, 7.45844623551049, 3.28459081620233, 
    3.4, 2.5
  )
) %>% 
  dplyr::filter(timepoint < 20)

data <- data %>% 
dplyr::bind_rows(
    data %>% 
  dplyr::filter(accession == "Col0") %>% 
  dplyr::mutate(
      timepoint = timepoint * time_stretch + time_shift,
      mean_cpm = mean_cpm * exp_stretch + exp_shift,
      accession = "Ro18"
    )
  ) 

data %>% 
  knitr::kable()

data %>% 
  ggplot() +
  aes(
    x = timepoint,
    y = mean_cpm,
    color = accession,
    shape = accession
  ) +
  geom_point() +
  geom_line(alpha = 0.5)
```

### Applying registration

```{r}
gg <- data %>% 
  dplyr::bind_rows(
    data %>% 
  dplyr::filter(accession == "Col0") %>% 
  dplyr::mutate(
      timepoint = timepoint * time_stretch + time_shift,
      mean_cpm = mean_cpm * exp_stretch + exp_shift,
      accession = "Col0 (Reg)"
    )
  ) %>%
  ggplot() +
  aes(
    x = timepoint,
    y = mean_cpm,
    color = accession,
    shape = accession
  ) +
  geom_point() +
  geom_line(alpha = 0.5)
```

```{r}
gg

gg + facet_wrap(~accession)
```

```{r}
data_after_reg <- data %>% filter(accession == "Ro18")
```

```{r}
do_splines <- function(degree = 3, df) {
  fit <- lm(formula = mean_cpm ~ splines::bs(timepoint, degree = degree, df = df), data = data_after_reg)
  # print(summary(fit))

  x_pred <- seq(min(data_after_reg$timepoint), max(data_after_reg$timepoint), len = 100)
  splines_df <- data.frame(
    timepoint = x_pred,
    mean_cpm = predict(fit, data.frame(timepoint = x_pred))
  )

  gg <- data_after_reg %>%
    ggplot() +
    aes(x = timepoint, y = mean_cpm) +
    geom_line(
      data = splines_df,
      color = "red"
    ) +
    geom_point() +
    labs(title = paste0("df: ", df, ", deg: ", degree))

  return(list(BIC = round(BIC(fit), 1), plot = gg))
}
```

```{r}
do_splines(degree = 3, df = 3)
do_splines(degree = 3, df = 4)
```

`degree + 1 <= df <= n - 2`

```{r residuals-loglik}
fit$residuals

# LogLik

logLik(fit)
AIC(fit)
```


### Data with noise 

```{r}
data
```
```{r}
set.seed(9)
data_with_noise <- data %>% 
  dplyr::mutate(mean_cpm = ifelse(accession == "Col0", mean_cpm + rnorm(4, sd = 2), mean_cpm))

data_with_noise %>% 
  ggplot() +
  aes(
    x = timepoint,
    y = mean_cpm,
    color = accession,
    shape = accession
  ) +
  geom_point() +
  geom_line(alpha = 0.5)
```

```{r}
gg <- data_with_noise %>% 
  dplyr::bind_rows(
    data_with_noise %>% 
  dplyr::filter(accession == "Col0") %>% 
  dplyr::mutate(
      timepoint = timepoint * time_stretch + time_shift,
      mean_cpm = mean_cpm * exp_stretch + exp_shift,
      accession = "Col0 (Reg)"
    )
  ) %>%
  ggplot() +
  aes(
    x = timepoint,
    y = mean_cpm,
    color = accession,
    shape = accession
  ) +
  geom_point() +
  geom_line(alpha = 0.5)

gg
```

```{r fit-noise}
data_after_reg_with_noise <- data_with_noise %>%
  dplyr::filter(accession == "Ro18") %>%
  dplyr::bind_rows(
    data_with_noise %>%
      dplyr::filter(accession == "Col0") %>%
      dplyr::mutate(
        timepoint = timepoint * time_stretch + time_shift,
        mean_cpm = mean_cpm * exp_stretch + exp_shift,
        accession = "Col0 (Reg)"
      )
  )

# fit <- lm(formula = mean_cpm ~ splines::bs(timepoint, degree = 2, knot = c(62)), data = data_after_reg)
fit <- lm(formula = mean_cpm ~ splines::bs(timepoint, degree = 3, df = 3), data = data_after_reg_with_noise)
summary(fit)
```


```{r gg-prediction-noise, warning=FALSE}
x_pred <- seq(min(data_after_reg_with_noise$timepoint), max(data_after_reg_with_noise$timepoint), len = 100)
splines_df = data.frame(
  timepoint = x_pred,
  mean_cpm = predict(fit, data.frame(timepoint = x_pred))
)
 
data_after_reg_with_noise %>% 
  ggplot() +
  aes(x = timepoint, y = mean_cpm) +
  geom_line(
    data = splines_df,
    color = "red"
  ) +
  geom_point()
```

#### Calculate AIC and BIC for non-registered data

```{r fit-separate}
fit_Ro18_sep <- lm(
  formula = mean_cpm ~ splines::bs(timepoint, degree = 3, df = 3), 
  data = data_with_noise %>% filter(accession == "Ro18")
)

fit_Col0_sep <- lm(
  formula = mean_cpm ~ splines::bs(timepoint, degree = 3, df = 3), 
  data = data_with_noise %>% filter(accession == "Col0")
)

data.frame(
  acession = c("Ro18", "Col0"),
  logLik = list(fit_Ro18_sep, fit_Col0_sep) %>% purrr::map(logLik) %>% unlist(),
  AIC =    list(fit_Ro18_sep, fit_Col0_sep) %>% purrr::map(AIC) %>% unlist(),
  BIC =    list(fit_Ro18_sep, fit_Col0_sep) %>% purrr::map(BIC) %>% unlist()
) %>% 
  knitr::kable()
```
