---
title: "Running registration using simulated data with repetition"
output:
  github_document:
    toc: true
---

This is a small experiment to test the standard deviation threshold for the simulated data with different number of timepoints.

## Set-up and load libraries

```{r setup, message=FALSE}
knitr::opts_chunk$set()

devtools::load_all()
library(dplyr)
library(data.table)
library(ggplot2)
library(cowplot)
library(ggpubr)
library(viridis)
library(ggridges)

theme_set(theme_bw())
```

```{r parallel-setup, eval=FALSE}
library(future)
future::plan(future::multisession, workers = 15)
```
## Function to simulate data

```{r fn-simulated-data}
#  Function to create data to register
simulate_data_to_reg <- function(gene_name, num_points, replicates = 1, replicate_sd = 1, method = c("theory", "fixed_var")) {
  method <- match.arg(method)

  timepoints_list <- seq(from = 10, by = 3, length.out = num_points)

  simulate_replicates_theory <- function(timepoint, gene_name, replicates, replicate_sd) {
    df <- data.frame(
      locus_name = gene_name,
      accession = "data_to_reg",
      timepoint = timepoint,
      mean_cpm = abs(rnorm(n = 1, sd = 10)) %>%
        purrr::map(
          ~ rnorm(mean = .x, n = replicates, sd = replicate_sd)
        ) %>%
        unlist(),
      group = as.factor(1:replicates)
    )

    return(df)
  }

  simulate_replicates_fixed_var <- function(timepoint, gene_name, replicates, replicate_sd) {
    df <- data.frame(
      locus_name = gene_name,
      accession = "data_to_reg",
      timepoint = timepoint,
      mean_cpm = abs(rnorm(n = 1, sd = 10)) %>%
        `+`(c(-replicate_sd, 0, replicate_sd)) %>%
        sample(replicates),
      group = as.factor(1:replicates)
    )

    return(df)
  }

  if (method == "theory") {
    df <- timepoints_list %>%
      purrr::map(
        ~ simulate_replicates_theory(.x, gene_name, replicates, replicate_sd)
      )
  } else if (method == "fixed_var") {
    if (replicates != 3) {
      stop("Replicates can only by 3 for this method.")
    }

    df <- timepoints_list %>%
      purrr::map(
        ~ simulate_replicates_fixed_var(.x, gene_name, replicates, replicate_sd)
      )
  }

  df <- df %>%
    purrr::reduce(dplyr::bind_rows)

  return(df)
}

# Function to create data fixed
simulate_data_fixed <- function(df_to_reg, time_stretch, time_shift, exp_stretch, exp_shift, num_points, exp_noise_sd) {
  df <- df_to_reg %>%
    dplyr::mutate(
      accession = "data_fixed",
      timepoint = timepoint * time_stretch + time_shift,
      mean_cpm = mean_cpm * exp_stretch + exp_shift + rnorm(n = num_points, sd = exp_noise_sd)
    )

  return(df)
}
```

```{r fn-undersample-data_timepoints-skip}
undersample_data_timepoints_skip <- function(data, keep_first_last_timepoint = TRUE, skip_timepoints_pattern = c(TRUE), columns_to_group = c("locus_name", "accession")) {
  if (keep_first_last_timepoint) {
    sliced_data <- data %>%
      dplyr::group_by(dplyr::across(columns_to_group)) %>%
      dplyr::arrange(timepoint) %>%
      dplyr::slice(c(1, seq(2, dplyr::n() - 1, 1)[skip_timepoints_pattern], dplyr::n())) %>%
      dplyr::ungroup()
  } else {
    sliced_data <- data %>%
      dplyr::group_by(dplyr::across(columns_to_group)) %>%
      dplyr::arrange(timepoint) %>%
      dplyr::slice(seq(1, dplyr::n(), 1)[skip_timepoints_pattern]) %>%
      dplyr::ungroup()
  }

  return(as.data.table(sliced_data))
}
```

### Test functions

```{r test-simulate-data-registered, eval=FALSE}
# Gene name list
list_letter <- LETTERS[seq(from = 1, to = 2)]
list_num <- seq(1, 3, 1)
gene_name <- paste0(list_letter, "_", list_num)

# Simulate registered data
data_to_reg_points <- gene_name %>%
  purrr::map(simulate_data_to_reg, num_points = 25, replicates = 3, replicate_sd = 4, method = "fixed_var") %>%
  purrr::reduce(rbind)
```

```{r test-simulate-data-plot, eval=FALSE}
data_to_reg_points %>%
  group_by(locus_name, accession, timepoint) %>%
  summarise(sd = sd(mean_cpm))

data_to_reg_points %>%
  ggplot() +
  aes(x = timepoint, y = mean_cpm, colour = group) +
  geom_point() +
  geom_line() +
  facet_wrap(accession ~ locus_name)
```


```{r simulate-data-fixed, eval=FALSE}
data_fixed <- simulate_data_fixed(
  data_to_reg_points,
  num_points = 25,
  time_stretch = 4,
  time_shift = 5,
  exp_stretch = 4,
  exp_shift = 5,
  exp_noise_sd = 2
)

simulated_data <- rbind(data_to_reg_points, data_fixed) %>%
  dplyr::mutate(
    accession = dplyr::case_when(
      accession == "data_to_reg" ~ "Col0",
      accession == "data_fixed" ~ "Ro18"
    )
  ) %>%
  as.data.table()

simulated_data %>% head()
```

```{r warning=FALSE, fig.width=15, fig.height=10, eval=FALSE}
simulated_data %>%
  ggplot() +
  aes(x = timepoint, y = mean_cpm, colour = group) +
  geom_point() +
  geom_line() +
  facet_wrap(accession ~ locus_name)

simulated_data %>%
  ggplot() +
  aes(x = timepoint, y = mean_cpm, colour = accession) +
  geom_point() +
  geom_line() +
  facet_wrap(group ~ locus_name, scales = "free_y")
```


```{r}
# simulated_data_mean <- simulated_data %>%
#   dplyr::group_by(locus_name, accession, timepoint) %>%
#   # dplyr::summarise(mean_cpm = mean(mean_cpm), n = n(), .groups = "drop") %>%
#   dplyr::summarise(mean_cpm = mean(mean_cpm), .groups = "drop") %>%
#   as.data.table()
#
# simulated_data_mean %>% head()
```

## Running simulations

```{r simulation-parameters}
# Simulation parameters
num_simulation <- 1:500
replicate_sd_list <- seq(0.25, 0.4, length.out = 16)
num_timepoints <- 10

# Simulated gene names
num_genes <- 10
list_letter <- LETTERS[1:2]
list_num <- 1:num_genes
gene_name <- paste0(list_letter, "_", list_num)
```

```{r running-multiple-simulation, warning=FALSE, message=FALSE, eval =FALSE}
results_list <- list()

for (simul in num_simulation) {
  for (rep_sd in replicate_sd_list) {
    data_to_reg_points <- gene_name %>%
      furrr::future_map(simulate_data_to_reg, num_points = num_timepoints, replicates = 3, replicate_sd = rep_sd, method = c("fixed_var")) %>%
      purrr::reduce(rbind)

    data_fixed <- simulate_data_fixed(
      data_to_reg_points,
      num_points = num_timepoints,
      time_stretch = 4,
      time_shift = 5,
      exp_stretch = 6,
      exp_shift = 4,
      exp_noise_sd = 2
    )

    simulated_data <- rbind(data_to_reg_points, data_fixed) %>%
      dplyr::mutate(
        accession = dplyr::case_when(
          accession == "data_to_reg" ~ "Col0",
          accession == "data_fixed" ~ "Ro18"
        )
      ) %>%
      as.data.table()

    simulated_data_mean <- simulated_data %>%
      dplyr::group_by(locus_name, accession, timepoint) %>%
      dplyr::summarise(mean_cpm = mean(mean_cpm), .groups = "drop") %>%
      as.data.table()

    # From the data, skip one timepoint
    skip_1_timepoint_all <- undersample_data_timepoints_skip(simulated_data, keep_first_last_timepoint = TRUE, skip_timepoints_pattern = c(FALSE, TRUE), columns_to_group = c("locus_name", "accession", "group"))
    skip_1_timepoint_mean <- undersample_data_timepoints_skip(simulated_data_mean, keep_first_last_timepoint = TRUE, skip_timepoints_pattern = c(FALSE, TRUE))

    # From the data, skip two timepoint
    skip_2_timepoint_all <- undersample_data_timepoints_skip(simulated_data, keep_first_last_timepoint = TRUE, skip_timepoints_pattern = c(FALSE, FALSE, TRUE), columns_to_group = c("locus_name", "accession", "group"))
    skip_2_timepoint_mean <- undersample_data_timepoints_skip(simulated_data_mean, keep_first_last_timepoint = TRUE, skip_timepoints_pattern = c(FALSE, FALSE, TRUE))

    # Save results
    sim_str <- paste0("sim_", stringr::str_pad(simul, width = 3, pad = "0"))
    rep_sd_str <- stringr::str_pad(rep_sd, width = 2, pad = "0")

    results_list[[paste(sim_str, "data_full_sd", rep_sd_str, sep = "_")]] <- list(
      all_data = simulated_data,
      mean_data = simulated_data_mean,
      sd = rep_sd,
      num_points = simulated_data_mean %>%
        dplyr::filter(accession == simulated_data_mean$accession[1]) %>%
        dplyr::pull(timepoint) %>% unique() %>% length(),
      num_simul = simul
    )

    results_list[[paste(sim_str, "data_skip1_sd", rep_sd_str, sep = "_")]] <- list(
      all_data = skip_1_timepoint_all,
      mean_data = skip_1_timepoint_mean,
      sd = rep_sd,
      num_points = skip_1_timepoint_mean %>%
        dplyr::filter(accession == simulated_data_mean$accession[1]) %>%
        dplyr::pull(timepoint) %>% unique() %>% length(),
      num_simul = simul
    )

    results_list[[paste(sim_str, "data_skip2_sd", rep_sd_str, sep = "_")]] <- list(
      all_data = skip_2_timepoint_all,
      mean_data = skip_2_timepoint_mean,
      sd = rep_sd,
      num_points = skip_2_timepoint_mean %>%
        dplyr::filter(accession == simulated_data_mean$accession[1]) %>%
        dplyr::pull(timepoint) %>% unique() %>% length(),
      num_simul = simul
    )
  }
}
```

```{r running-multiple-simulation-b, warning=FALSE, eval=FALSE}
total_sims <- length(num_simulation)
sim_count <- numeric(length = total_sims)

reg_with_full_simulated_data <- results_list %>%
  furrr::future_map(
    function(x) {
      # Check simulation processing progress
      num_simul <- x$num_simul
      if (!(num_simul %in% sim_count)) {
        sim_count[num_simul] <<- num_simul
        message("[", Sys.time(), "] ", "Processing simulation ", num_simul, "/", total_sims, "...")
      }

      # Register data
      if (x$num_points == 10) {
        results <- GREAT::scale_and_register_data(
          x$mean_data %>% dplyr::mutate(tissue = "apex"),
          x$all_data %>% dplyr::mutate(tissue = "apex"),
          stretches =  c(4, 3, 2, 1),
          shift_extreme = 5,
          num_shifts = 21,
          min_num_overlapping_points = 4,
          initial_rescale = FALSE,
          do_rescale = TRUE,
          testing = FALSE,
          accession_data_to_transform = "Col0",
          accession_data_fix = "Ro18",
          data_to_transform_time_added = 40,
          data_fix_time_added = 40
        ) %>%
          suppressMessages()
      } else {
        results <- GREAT::scale_and_register_data(
          x$mean_data %>% dplyr::mutate(tissue = "apex"),
          x$all_data %>% dplyr::mutate(tissue = "apex"),
          stretches =  c(4, 3, 2, 1),
          shift_extreme = 5,
          num_shifts = 21,
          min_num_overlapping_points = 2,
          initial_rescale = FALSE,
          do_rescale = TRUE,
          testing = FALSE,
          accession_data_to_transform = "Col0",
          accession_data_fix = "Ro18",
          data_to_transform_time_added = 40,
          data_fix_time_added = 40
        ) %>%
          suppressMessages()
      }

      results[["simul"]] <- x$num_simul
      results[["sd"]] <- x$sd
      results[["num_points"]] <- x$num_points

      return(results)
    }
  )
```

```{r save-results, include=FALSE, eval=FALSE}
timestamp <- Sys.time() %>%
  stringr::str_replace_all(" ", "_") %>%
  stringr::str_replace_all(":", "-")

saveRDS(results_list, file = paste0("~/Downloads/results_list_", timestamp, ".rds"))
saveRDS(reg_with_full_simulated_data, file = paste0("~/Downloads/reg_with_full_simulated_data_", timestamp, ".rds"))
```
### Analysing data

```{r read-data, include=FALSE}
# Reading data from simulation
# reg_with_full_simulated_data <- readRDS("~/Downloads/reg_with_full_simulated_data_2021-10-11_15-57-54.rds")
# reg_with_full_simulated_data_100_simulation <- readRDS("~/Downloads/reg_with_full_simulated_data_11Oct.rds")
table_100simul <- data.table::fread(here::here("dev/table_100simul_reg_with_full_simulated_data_11Oct.csv")) %>% 
  dplyr::mutate(
    num_timepoints = as.factor(num_timepoints),
    SD = as.factor(SD)
  )
table <- data.table::fread(here::here("dev/table_reg_with_full_simulated_data_2021-10-11_15-57-54.csv")) %>% 
  dplyr::mutate(
    num_timepoints = as.factor(num_timepoints),
    SD = as.factor(SD)
  )

result <- readRDS("~/Downloads/results_list_2021-10-11_15-57-54.rds")

result$sim_001_data_full_sd_0.25 %>% head(5)
```

```{r simulation-results, eval=FALSE}
table <- reg_with_full_simulated_data %>%
  purrr::map(
    ~ data.frame(
      num_simul = .x$simul,
      SD = .x$sd,
      num_timepoints = .x$num_points,
      `registered_genes_total10` = .x$model_comparison_dt %>%
        dplyr::filter(BIC_registered_is_better == TRUE) %>%
        dplyr::nrow()
    )
  ) %>%
  purrr::reduce(dplyr::bind_rows) %>%
  dplyr::mutate(
    num_timepoints = as.factor(num_timepoints),
    SD = as.factor(SD)
  )


table_100simul <- reg_with_full_simulated_data_100_simulation %>%
  purrr::map(
    ~ data.frame(
      num_simul = .x$simul,
      SD = .x$sd,
      num_timepoints = .x$num_points,
      `registered_genes_total10` = .x$model_comparison_dt %>%
        dplyr::filter(BIC_registered_is_better == TRUE) %>%
        nrow()
    )
  ) %>%
  purrr::reduce(dplyr::bind_rows) %>%
  dplyr::mutate(
    num_timepoints = as.factor(num_timepoints),
    SD = as.factor(SD)
  )
```


```{r fn-get-bar-plot}
get_bar_plot <- function(data) {
  plot <- data %>%
    dplyr::group_by(SD, num_timepoints) %>%
    dplyr::summarise(
      dplyr::across(
        .cols = registered_genes_total10,
        list(mean = mean, sd = sd)
      ),
      .groups = "drop"
    ) %>%
    dplyr::rename(mean_length = registered_genes_total10_mean, sd_length = registered_genes_total10_sd) %>%
    # Plot
    ggplot() +
    aes(y = mean_length, x = as.factor(num_timepoints)) +
    geom_bar(
      stat = "identity",
      fill = "#8da0cb",
      position = position_dodge()
    ) +
    geom_errorbar(
      aes(
        ymin = mean_length - sd_length,
        ymax = mean_length + sd_length
      ),
      width = .2,
      position = position_dodge(.9)
    ) +
    labs(
      x = "Number timepoints",
      y = "Registered genes",
      title = "Total registered genes"
    ) +
    facet_wrap(~SD)

  return(plot)
}
```

## Barplot with standard error of simulated data for n=100

```{r plot-table-100-simulation, fig.width=10, fig.height=10}
table_100simul %>%
  get_bar_plot()
```

## Barplot with standard error of simulated data for n=500

```{r plot-table-500-simulation, fig.width=10, fig.height=10}
table %>%
  get_bar_plot()
```

## Boxplot of simulated data

```{r boxplot-with-jitter, fig.width=10, fig.height=12}
table %>%
  ggplot(aes(y = registered_genes_total10, x = num_timepoints, fill = num_timepoints)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha = 0.6) +
  geom_jitter(color = "black", size = 0.4, alpha = 0.1) +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 11)
  ) +
  scale_y_continuous(breaks = seq(1, 10, by = 1)) +
  labs(x = "Number timepoints", y = "Number of registered genes") +
  facet_wrap(~SD, ncol = 4)
```

```{r percentage-registered-genes-plot, fig.width=10, fig.height=4}
table %>%
  dplyr::mutate(
    registered_genes_total10 = as.factor(registered_genes_total10),
    SD = as.numeric(as.character(SD))
  ) %>%
  dplyr::filter(
    SD >= 0.3,
    SD <= 0.4
  ) %>%
  dplyr::group_by(SD, num_timepoints, registered_genes_total10) %>%
  dplyr::summarise(
    count = n(),
    percentage = n() / 500 * 100,
    .groups = "drop"
  ) %>%
  ggplot() +
  aes(x = SD, y = percentage, group = registered_genes_total10, color = registered_genes_total10) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = 0.35, linetype = "dashed") +
  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
  labs(
    x = "Standard deviation",
    y = "Percentage",
    color = "Number of registered genes"
  ) +
  facet_wrap(
    ~num_timepoints,
    labeller = labeller(
      num_timepoints =
        c(
          "4" = "Number of points: 4",
          "6" = "Number of points: 6",
          "10" = "Number of points: 8"
        )
    )
  )
```

## ANOVA (ANalysis Of VAriance) test

- Null hypothesis ($H_{0}$): $\mu_{\sigma_i, \forall n} = \mu_{\sigma_j, \forall n} = \dots = \mu_{\sigma_k, \forall n}$.

That is, for $\sigma_{i}$ all groups with different time points ($n$) have same mean of number registered genes.

- Alternative hypothesis ($H_{1}$): not all samples have the same mean of number registered genes.

The null hypothesis is tested by comparing variances.


```{r anova-test}
aov_results <- aov(
  formula = registered_genes_total10 ~ SD,
  data = table
)
```

```{r anova-results}
summary(aov_results)
```

The p-value is far below 0.05, so the null is rejected, and it can be concluded there are differences between mean of number of registered genes between different groups with different SD.

```{r TukeyHSD-results}
TukeyHSD_result <- TukeyHSD(aov_results)

TukeyHSD_result$SD %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column() %>% 
  tidyr::separate(rowname, into = c("sd_a", "sd_b"), sep = "-") %>% 
  dplyr::mutate(
    across(c(sd_a, sd_b), as.numeric)
  ) %>% 
  dplyr::group_by(sd_b) %>% 
  dplyr::summarise(
    good = sum(`p adj` >= 0.05),
    total = dplyr::n()
  ) %>% 
  dplyr::filter(good == total)
```


```{r}
TukeyHSD_result$SD %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column() %>% 
  tidyr::separate(rowname, into = c("sd_a", "sd_b"), sep = "-") %>% 
  dplyr::mutate(
    across(c(sd_a, sd_b), as.numeric)
) %>% 
  dplyr::filter(sd_b >= 0.28) %>% 
  dplyr::select(sd_a, sd_b, `p adj`) %>% 
  knitr::kable()
```

