---
title: "Gene registration -- running all functions for *B.rapa* key floral genes"
output: 
  github_document:
    toc: true
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(genereg)
library(ggplot2)
library(data.table)

source("scripts/chiifu_v3_scripts/unused_functions.R")
```

# Apply all of the registration function to *B.rapa* key floral genes

## Background

This short report is to try to run all of the registration functions, with the purpose to get/reproduce the same results in Alex's paper. 

```r
do.initial.rescale <- 'rescale' # should be 'rescale' if want to use scaled df for registration, rather than mean.df
do.register.rescale <- 'rescale' 
outdir.string <- 'TESTING_rescale_as_register___shuffled_g_v4__'
```

```r
# calculate the scores associated with each candidate shift
# Arabidopsis observations = 7d -> 16d = 0d -> 9d
# Ro18 observations = 11d -> 35d = 0d -> 24d
# therefore if same start and end points stretch = 24 / 9

# lots of spurious overlaps detected when too extreme shifts allowed
stretch = c(2, 1.5, 1) 
min.num.overlapping.points = 4
shift.extreme  = 4 
transformed.timecourse = 'Col0' 
num.shuffled <- 1 
jobNum <- 1
```

## Run scripts

```r
# setup flags for rescaling options
if (do.initial.rescale=='rescale') {
  initial.rescale <- TRUE
} else {
  initial.rescale <- FALSE
}
if (do.register.rescale=='rescale') {
  should.rescale <- TRUE
} else {
  should.rescale <- FALSE
}

if (initial.rescale==TRUE) {
  print('********************')
  print('will rescale the data prior to registering, and register using this rescaled mean data!')
  print('********************')
}
if (should.rescale==TRUE){
  print('********************')
  print('will rescale the data when deciding optimal registration!')
  print('********************')
} 
```

```r
# directories to save graphs to
real.data.graph.dir <- paste0('graphs/gene_registration/', outdir.string, '_real_data/')
# shuffled.data.graph.dir <- paste0('graphs/gene_registration/', outdir.string, '_shuffled_data/job_', jobNum, '/')
# directories to save real and shuffled expression data to
real.expression.dir <- paste0('intermediate_data/gene_registration/', outdir.string, '_real_data/gene_expression/job_', jobNum, '/')
shuffled.expression.dir <- paste0('intermediate_data/gene_registration/', outdir.string, '_shuffled_data/gene_expression/')
# directories to save real and shuffled distance data to
real.distance.dir <- paste0('intermediate_data/gene_registration/', outdir.string, '_real_distance/job_', jobNum, '/')
shuffled.distance.dir <- paste0('intermediate_data/gene_registration/', outdir.string, '_shuffled_distance/')
  

# somewhere to store the data.tables and graphs
if (!(dir.exists(real.expression.dir))) {
    dir.create(real.expression.dir, recursive=T)
    dir.create(shuffled.expression.dir, recursive=T)
    
    dir.create(real.distance.dir, recursive=T)
    dir.create(shuffled.distance.dir, recursive=T)

    dir.create(real.data.graph.dir, recursive=T)
    dir.create(shuffled.data.graph.dir, recursive=T)
}
```

```r
## GET THE RAW DATA. MEAN EXPRESSION OF BIOLOGICAL REPS. TREAT BRASSICA GENES INDIVIDUALLY (DON'T SUM THEM).
# load the data expression data. Consider the brassica genes individually (don't sum)
L <- load_mean.df() # sumBrassica copy flag is within load_mean.df()
mean.df <- L[[1]] # will compare Col0 vs Ro18 based on accession column
all.data.df <- L[[2]]
```

```r
## PREPARE, AND REGISTER AND SCALE THE DATA
O <- prepare_scaled_and_registered_data(mean.df, all.data.df,
  stretch = stretch, initial.rescale, should.rescale, min.num.overlapping.points,
  shift.extreme, transformed.timecourse
)

mean.df <- O[["mean.df"]] # mean.df is unchanged
mean.df.sc <- O[["mean.df.sc"]] # mean.df.sc : data is scaled(center=T, scale=T)
imputed.mean.df <- O[["imputed.mean.df"]] # imputed.mean.df is registered data, Col0 values imputed to make a single timepoint.
all.shifts <- O[["all.shifts"]] # all.shifts is data.table of score for each shift for each gene.
model.comparison <- O[["model.comparison"]]

# sanity test plot
# ggplot(imputed.mean.df[imputed.mean.df$locus_name=='BRAA02G015410.3C',],
#        aes(x=shifted.time, y=mean.cpm, color=accession))+
#   geom_point()+
#   geom_line()

#### CALCULATE THE DISTANCES BETWEEN THE SAMPLES ####
O <- calculate_between_sample_distance(mean.df, mean.df.sc, imputed.mean.df)
D.mean <- O[["D.mean"]]
D.scaled <- O[["D.scaled"]]
D.registered <- O[["D.registered"]]
D.scaled.onlyNR <- O[["D.scaled.onlyNR"]]
D.scaled.onlyR <- O[["D.scaled.onlyR"]]
D.registered.onlyR <- O[["D.registered.onlyR"]]
```

```r
saveRDS(mean.df, file = paste0(real.expression.dir, "mean.df.rds"))
saveRDS(mean.df.sc, file = paste0(real.expression.dir, "mean.df.sc.rds"))
saveRDS(imputed.mean.df, file = paste0(real.expression.dir, "imputed.mean.df.rds"))
saveRDS(all.shifts, file = paste0(real.expression.dir, "all.shifts.rds"))
saveRDS(model.comparison, file = paste0(real.expression.dir, "model.comparison.rds"))

# save the distances calculated
saveRDS(D.mean, file = paste0(real.distance.dir, "D.mean.rds"))
saveRDS(D.scaled, file = paste0(real.distance.dir, "D.scaled.rds"))
saveRDS(D.registered, file = paste0(real.distance.dir, "D.registered.rds"))
saveRDS(D.scaled.onlyNR, file = paste0(real.distance.dir, "D.scaled.onlyNR.rds"))
saveRDS(D.scaled.onlyR, file = paste0(real.distance.dir, "D.scaled.onlyR.rds"))
saveRDS(D.registered.onlyR, file = paste0(real.distance.dir, "D.registered.onlyR.rds"))
```

## Get the saved RDS data

```{r}
mean.df <- readRDS("intermediate_data/gene_registration/TESTING_rescale_as_register___shuffled_g_v4___real_data/gene_expression/job_1/mean.df.rds")
imputed.mean.df <- readRDS("intermediate_data/gene_registration/TESTING_rescale_as_register___shuffled_g_v4___real_data/gene_expression/job_1/imputed.mean.df.rds")
model.comparison <- readRDS("intermediate_data/gene_registration/TESTING_rescale_as_register___shuffled_g_v4___real_data/gene_expression/job_1/model.comparison.rds")
```



```{r}
mean.df %>% 
  head(10) %>% 
  knitr::kable()
```


```{r}
imputed.mean.df %>% 
  head(10) %>% 
  knitr::kable()
```

```{r}
model.comparison %>% 
  head(10) %>% 
  knitr::kable()
```



### Take the reference

This reference table contains all of the name of key floral genes. 

```{r}
id_table <- readRDS("reference_data/ID_TABLE_brapa-v3.rds")

id_table_unique_key_floral <- unique(id_table[, c('locus_name', 'symbol', 'CDS.model')]) %>% 
  dplyr::filter(!is.na(symbol)) %>% 
  dplyr::filter(symbol %in% c("SOC1", "LFY", "AP1", "AP3", "AGL24"))
```

```{r}
id_table_unique_key_floral %>% 
  knitr::kable()
```

```{r}
# get list of CDS.model
cds_model_wanted <- id_table_unique_key_floral %>% 
  dplyr::pull(CDS.model)
```

### Get the registered key floral genes

```{r}
imputed.mean.df_key_floral <- imputed.mean.df %>% 
  dplyr::filter(locus_name %in% cds_model_wanted)
```


```{r}
# left join to get the symbol genes (e.g. AP3 etc) to the main df
imputed.mean.df_wanted <- imputed.mean.df_key_floral %>%
  dplyr::left_join(id_table_unique_key_floral, by = c("locus_name" = "CDS.model")) %>% 
  dplyr::select(-c(locus_name.y)) %>%
  dplyr::rename(locus_name = symbol, bra_gene = locus_name)
```


### Plot the registered genes

```{r warning=FALSE}
plot_registered_GoIs_for_comparible_timepoints(imputed.mean.df_wanted)
```

Figure in Ales's paper:

```{r alexfig, fig.width=75, fig.height=50, echo=FALSE}
library(png)
library(grid)

grid.raster(readPNG("figure/alexs_fig.png"))
```


### Debug (what happened)

#### Missing "CDS.model"?

```{r}
raw <- readRDS("final_data/rds/ro18_chiifu_apex.rds")
```

```{r}
raw %>% 
  head(5) %>% 
  knitr::kable()
```

```{r}
id_table_unique_key_floral %>% 
  knitr::kable()
```


```{r}
cds_model_raw <-raw %>% 
  dplyr::filter(CDS.model %in% cds_model_wanted) %>%
  dplyr::pull(CDS.model) %>% 
  unique()

cds_model_raw
```

```{r}
cds_model_wanted
```


```{r}
dplyr::setdiff(cds_model_wanted, cds_model_raw)
```

```{r}
id_table_unique_key_floral %>% 
  dplyr::filter(CDS.model == dplyr::setdiff(cds_model_wanted, cds_model_raw)) %>% 
  knitr::kable()
```

#### Do they actually have the same optimal parameters?

```{r}
model.comparison %>% 
  dplyr::filter(gene %in% cds_model_wanted) %>% 
  dplyr::select(gene, stretch, shift) %>% 
  dplyr::left_join(id_table_unique_key_floral, by = c("gene" = "CDS.model")) %>% 
  dplyr::select(symbol, gene, stretch, shift) %>% 
  dplyr::arrange(symbol) %>% 
  knitr::kable()
```

```{r}
seq(-4, 4, length.out = 10)
```

The paramater used by alex 

```{r}
seq(-4, 4, length.out = 13)
```

















